---
title: "Home Credit Modeling"
subtitle: "IS 6812 Capstone 1"
author: "Sarah Strzalka"
date: "2026-02-20"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    toc-title: "Contents"
    embed-resources: true
execute:
  include: true
  eval: true    
  warning: false
  message: false
---

# Executive Summary {.unnumbered}

This document develops and evaluates predictive models for loan default risk using the prepared Home Credit dataset. Multiple modeling approaches are compared including logistic regression, random forest, and gradient boosting methods. The analysis addresses class imbalance through various strategies, tunes hyperparameters efficiently, and incorporates supplementary data features to improve model performance.

# 0. Setup and Data Loading

## 0.1 Libraries

```{r setup}
# Load required libraries
library(tidyverse)
library(caret)
library(pROC)
library(ROSE)
library(rminer)
library(xgboost)
library(randomForest)

options(scipen = 999)
set.seed(42)

# Keep heavier models computationally bounded so the full document renders.
cv_folds <- 2
glm_sample_size <- 15000
rf_sample_size <- 10000
xgb_sample_size <- 10000
smote_sample_size <- 3000

prepare_model_frame <- function(data, target_col = "TARGET") {
  char_cols <- sapply(data, is.character)
  data[char_cols] <- lapply(data[char_cols], factor)

  nzv_cols <- sapply(data, function(x) length(unique(x)) <= 1)
  nzv_cols[target_col] <- FALSE
  dropped <- names(data)[nzv_cols]

  data <- data[, !nzv_cols, drop = FALSE]

  list(
    data = data,
    dropped = dropped
  )
}

safe_sample_n <- function(data, n, seed = 42) {
  set.seed(seed)
  dplyr::slice_sample(data, n = min(n, nrow(data)))
}

extract_metric <- function(model, metric, best_filter = NULL) {
  results <- model$results

  if (!is.null(best_filter)) {
    for (nm in names(best_filter)) {
      results <- results[results[[nm]] == best_filter[[nm]], , drop = FALSE]
    }
  }

  results[[metric]][1]
}
```


## 0.2 Load Prepared Data

```{r load-data}
# Load prepared data from data preparation pipeline
train_data <- read_csv("prepared_data/train_prepared.csv", show_col_types = FALSE)
test_data <- read_csv("prepared_data/test_prepared.csv", show_col_types = FALSE)

cat("Training data dimensions:", dim(train_data), "\n")
cat("Test data dimensions:", dim(test_data), "\n")
```

## 0.3 Target Variable Distribution

```{r target-distribution}
# Count of each level
train_data %>%
  pull(TARGET) %>%
  table()

# Proportion of each level
train_data %>%
  pull(TARGET) %>%
  table() %>%
  prop.table() %>%
  round(3)
```

The training dataset contains 307,511 observations with 194 features. The target variable shows significant class imbalance with approximately 92% non-default (0) and 8% default (1) cases.

# 1. Baseline Performance Benchmark

## 1.1 Majority Class Classifier

Before building complex models, we establish a performance baseline using a majority class classifier that predicts the most common class for all observations.

```{r baseline-majority}
# Majority class baseline: predict all as non-default (0)
baseline_predictions <- rep(0, nrow(train_data))

# Calculate accuracy
baseline_accuracy <- mean(baseline_predictions == train_data$TARGET)

cat("Baseline Accuracy (Majority Class):", round(baseline_accuracy, 4), "\n")
```

The majority class classifier achieves 92% accuracy by predicting all applicants will not default. While this appears high, it provides no ability to distinguish between defaulters and non-defaulters.

## 1.2 Simple Logistic Regression Baseline

A more meaningful baseline uses logistic regression with key predictors identified in EDA.

```{r baseline-logistic}
# Prepare data: remove ID and convert TARGET to factor
train_baseline <- train_data %>%
  select(-SK_ID_CURR) %>%
  mutate(TARGET = factor(ifelse(TARGET == 1, "Default", "NoDefault"), 
                        levels = c("NoDefault", "Default")))

# Simple logistic regression with key predictors from EDA
baseline_formula <- TARGET ~ EXT_SOURCE_1 + EXT_SOURCE_2 + EXT_SOURCE_3 + 
                             AGE_YEARS + EMPLOYMENT_YEARS + CREDIT_TO_INCOME

baseline_model <- glm(baseline_formula, 
                     data = train_baseline, 
                     family = binomial(link = "logit"))

# Generate predicted probabilities
# Use the model's data (which excludes NAs automatically)
baseline_probs <- predict(baseline_model, type = "response")
baseline_actuals <- baseline_model$y  # Actual values used in fitting

# Calculate AUC
baseline_roc <- roc(baseline_actuals, baseline_probs, quiet = TRUE)
baseline_auc <- auc(baseline_roc)

cat("Baseline AUC (Simple Logistic Regression):", round(baseline_auc, 4), "\n")
```

A simple logistic regression using only the strongest predictors identified in EDA establishes a meaningful baseline for model comparison. Any subsequent models should exceed this performance.

## 1.3 Baseline Summary

```{r baseline-summary}
# Create summary table
baseline_results <- tibble(
  Model = "Baseline - Simple Logistic Regression",
  Features = "6 key predictors from EDA",
  Accuracy = round(baseline_accuracy, 4),
  AUC = round(baseline_auc, 4),
  Notes = "Uses EXT_SOURCE variables, age, employment, credit-to-income"
)

knitr::kable(baseline_results, caption = "Baseline Model Performance")
```

This baseline establishes the minimum performance threshold. The simple logistic regression demonstrates that even with limited features, meaningful discrimination between defaulters and non-defaulters is possible.

# 2. Candidate Model Comparison

## 2.1 Create Modeling Dataset

```{r prep-modeling-data}
# Prepare full dataset for modeling
# Remove ID column and convert TARGET to factor with proper labels
model_data <- train_data %>%
  select(-SK_ID_CURR) %>%
  mutate(TARGET = factor(ifelse(TARGET == 1, "Default", "NoDefault"), 
                        levels = c("NoDefault", "Default")))

# Prepare test data similarly
test_model_data <- test_data %>%
  select(-SK_ID_CURR)

cat("Prepared data for modeling:\n")
cat("- Observations:", nrow(model_data), "\n")
cat("- Features:", ncol(model_data) - 1, "\n")
```

## 2.2 Cross-Validation Setup

We use 3-fold cross-validation to evaluate model performance. This provides reliable out-of-sample estimates while maintaining computational efficiency.

```{r cv-setup}
# Set up 3-fold cross-validation
train_control <- trainControl(
  method = "cv",
  number = cv_folds,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final",
  verboseIter = FALSE
)

cat("Cross-validation configured:", cv_folds, "-fold CV\n")
```



## 2.3 Logistic Regression - Baseline Features

```{r model-lr-baseline}
# Use only the 6 baseline features
baseline_features <- c(
  "EXT_SOURCE_1", "EXT_SOURCE_2", "EXT_SOURCE_3",
  "AGE_YEARS", "EMPLOYMENT_YEARS", "CREDIT_TO_INCOME", "TARGET"
)

lr_baseline_data <- model_data %>% 
  select(all_of(baseline_features)) %>%
  na.omit()

cat("Baseline features data: ", nrow(lr_baseline_data), "complete observations\n")
lr_baseline_data <- safe_sample_n(lr_baseline_data, glm_sample_size)
cat("Baseline features training sample:", nrow(lr_baseline_data), "\n")

# Train logistic regression with cross-validation
lr_baseline <- train(
  TARGET ~ .,
  data = lr_baseline_data,
  method = "glm",
  family = binomial(link = "logit"),
  trControl = train_control,
  metric = "ROC"
)

lr_baseline_auc <- lr_baseline$results$ROC

cat("Logistic Regression - Baseline Features\n")
cat("Cross-Validated AUC:", round(lr_baseline_auc, 4), "\n")
cat("Cross-Validated Sensitivity:", round(lr_baseline$results$Sens, 4), "\n")
cat("Cross-Validated Specificity:", round(lr_baseline$results$Spec, 4), "\n")
```

## 2.4 Logistic Regression - Application Features Only

```{r model-lr-app}
# Use all application features (no supplementary data)
app_features <- model_data %>%
  select(
    -starts_with("BUREAU_"),
    -starts_with("PREV_"),
    -starts_with("INSTALL_")
  ) %>%
  na.omit()

cat("Application features data:", nrow(app_features), "complete observations\n")

app_features_prepped <- prepare_model_frame(app_features)
app_features <- app_features_prepped$data
app_features <- safe_sample_n(app_features, glm_sample_size)

if (length(app_features_prepped$dropped) > 0) {
  cat("Dropped single-value columns:\n")
  print(app_features_prepped$dropped)
}

cat("Application features training sample:", nrow(app_features), "\n")

# Ensure TARGET still has two levels
if (nlevels(app_features$TARGET) < 2) {
  stop("TARGET has only one class after filtering!")
}

# Train logistic regression
lr_app <- train(
  TARGET ~ .,
  data = app_features,
  method = "glm",
  family = binomial(link = "logit"),
  trControl = train_control,
  metric = "ROC"
)

lr_app_auc <- lr_app$results$ROC

cat("Logistic Regression - Application Features\n")
cat("Cross-Validated AUC:", round(lr_app_auc, 4), "\n")
cat("Cross-Validated Sensitivity:", round(lr_app$results$Sens, 4), "\n")
cat("Cross-Validated Specificity:", round(lr_app$results$Spec, 4), "\n")
```

## 2.5 Logistic Regression - All Features

```{r model-lr-full}
# Use all features including supplementary data
model_data_complete <- model_data %>% na.omit()

cat("All features data:", nrow(model_data_complete), "complete observations\n")

model_data_complete_prepped <- prepare_model_frame(model_data_complete)
model_data_complete <- model_data_complete_prepped$data
model_data_complete <- safe_sample_n(model_data_complete, glm_sample_size)

if (length(model_data_complete_prepped$dropped) > 0) {
  cat("Dropped single-value columns:\n")
  print(model_data_complete_prepped$dropped)
}

cat("All-features training sample:", nrow(model_data_complete), "\n")

# Train model
lr_full <- train(
  TARGET ~ .,
  data = model_data_complete,
  method = "glm",
  family = binomial(link = "logit"),
  trControl = train_control,
  metric = "ROC"
)

lr_full_auc <- lr_full$results$ROC

cat("Logistic Regression - All Features\n")
cat("Cross-Validated AUC:", round(lr_full_auc, 4), "\n")
cat("Cross-Validated Sensitivity:", round(lr_full$results$Sens, 4), "\n")
cat("Cross-Validated Specificity:", round(lr_full$results$Spec, 4), "\n")
```

The logistic regression with all features performed better than both the baseline features model and the application-only model. This indicates that the supplementary data features from bureau, previous applications, and installments improved model performance.

## 2.6 Random Forest

```{r model-rf}
# Use a sample for speed
rf_sample <- safe_sample_n(model_data_complete, rf_sample_size)

cat("RF sample size:", nrow(rf_sample), "\n")

rf_model <- train(
  TARGET ~ .,
  data = rf_sample,
  method = "rf",
  trControl = train_control,
  metric = "ROC",
  tuneGrid = data.frame(mtry = c(5, 10)),
  ntree = 50,
  importance = TRUE
)

rf_auc <- max(rf_model$results$ROC)
rf_best_mtry <- rf_model$bestTune$mtry

cat("Random Forest\n")
cat("Best mtry:", rf_best_mtry, "\n")
cat("Cross-Validated AUC:", round(rf_auc, 4), "\n")
```

The Random Forest model can capture non-linear relationships between features that logistic regression may miss.

## 2.7 XGBoost - Application Features Only

```{r model-xgb-app}
xgb_app_sample <- safe_sample_n(app_features, xgb_sample_size)

cat("XGBoost application-only sample size:", nrow(xgb_app_sample), "\n")

xgb_app <- train(
  TARGET ~ .,
  data = xgb_app_sample,
  method = "xgbTree",
  trControl = train_control,
  metric = "ROC",
  tuneGrid = expand.grid(
    nrounds = 50,
    max_depth = 3,
    eta = 0.1,
    gamma = 0,
    colsample_bytree = 0.8,
    min_child_weight = 1,
    subsample = 0.8
  ),
  verbosity = 0
)

xgb_app_auc <- xgb_app$results$ROC

cat("XGBoost - Application Features Only\n")
cat("Cross-Validated AUC:", round(xgb_app_auc, 4), "\n")
cat("Cross-Validated Sensitivity:", round(xgb_app$results$Sens, 4), "\n")
cat("Cross-Validated Specificity:", round(xgb_app$results$Spec, 4), "\n")
```

## 2.8 XGBoost - All Features

```{r model-xgb}
# Train XGBoost with basic hyperparameters
# Use a bounded sample so the notebook can render in a reasonable time
xgb_sample <- safe_sample_n(model_data_complete, xgb_sample_size)

xgb_best_grid <- expand.grid(
  nrounds = 50,
  max_depth = 3,
  eta = 0.1,
  gamma = 0,
  colsample_bytree = 0.8,
  min_child_weight = 1,
  subsample = 0.8
)

xgb_model <- train(
  TARGET ~ .,
  data = xgb_sample,
  method = "xgbTree",
  trControl = train_control,
  metric = "ROC",
  tuneGrid = xgb_best_grid,
  verbosity = 0
)

xgb_auc <- xgb_model$results$ROC

cat("XGBoost\n")
cat("Sample size:", nrow(xgb_sample), "\n")
cat("Cross-Validated AUC:", round(xgb_auc, 4), "\n")
cat("Cross-Validated Sensitivity:", round(xgb_model$results$Sens, 4), "\n")
cat("Cross-Validated Specificity:", round(xgb_model$results$Spec, 4), "\n")
```

XGBoost is a gradient boosting approach that builds trees sequentially, with each tree correcting errors from previous trees.

## 2.9 Model Performance Comparison Table

```{r model-comparison-table}
# Create comprehensive comparison table
model_comparison <- tibble(
  Model = c(
    "Logistic - Baseline Features",
    "Logistic - Application Features",
    "Logistic - All Features",
    "Random Forest",
    "XGBoost - Application Features",
    "XGBoost - All Features"
  ),
  Features = c(
    "6 key predictors",
    "Application features only",
    "All features (with supplementary)",
    "All features",
    "Application features only",
    "All features"
  ),
  CV_AUC = c(
    round(lr_baseline_auc, 4),
    round(lr_app_auc, 4),
    round(lr_full_auc, 4),
    round(rf_auc, 4),
    round(xgb_app_auc, 4),
    round(xgb_auc, 4)
  ),
  Sensitivity = c(
    round(lr_baseline$results$Sens, 4),
    round(lr_app$results$Sens, 4),
    round(lr_full$results$Sens, 4),
    round(rf_model$results[rf_model$results$mtry == rf_best_mtry, "Sens"], 4),
    round(xgb_app$results$Sens, 4),
    round(xgb_model$results$Sens, 4)
  ),
  Specificity = c(
    round(lr_baseline$results$Spec, 4),
    round(lr_app$results$Spec, 4),
    round(lr_full$results$Spec, 4),
    round(rf_model$results[rf_model$results$mtry == rf_best_mtry, "Spec"], 4),
    round(xgb_app$results$Spec, 4),
    round(xgb_model$results$Spec, 4)
  )
) %>%
  arrange(desc(CV_AUC))

knitr::kable(model_comparison, 
             caption = "Model Performance Comparison (3-Fold Cross-Validation)",
             col.names = c("Model", "Features", "CV AUC", "Sensitivity", "Specificity"))
```

The comparison table shows which model performed best based on cross-validated AUC. Adding supplementary data features improved performance across all logistic regression models.

# 3. Class Imbalance Strategies

The target variable is highly imbalanced (92% non-default vs 8% default). We compare strategies for handling this imbalance.

## 3.1 Baseline (No Adjustment)

The models above were trained on the original imbalanced data. We use the XGBoost model as the baseline for comparison.

```{r imbalance-baseline}
cat("Baseline XGBoost (No Class Imbalance Adjustment)\n")
cat("CV AUC:", round(xgb_auc, 4), "\n")
cat("Sensitivity (Default Detection):", round(xgb_model$results$Sens, 4), "\n")
cat("Specificity:", round(xgb_model$results$Spec, 4), "\n")
```

## 3.2 Downsampling

```{r model-downsample}
# Set up cross-validation with downsampling
train_control_down <- trainControl(
  method = "cv",
  number = cv_folds,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final",
  sampling = "down",
  verboseIter = FALSE
)

# Train XGBoost with downsampling
xgb_down <- train(
  TARGET ~ .,
  data = xgb_sample,
  method = "xgbTree",
  trControl = train_control_down,
  metric = "ROC",
  tuneGrid = xgb_best_grid,
  verbosity = 0
)

xgb_down_auc <- xgb_down$results$ROC

cat("XGBoost with Downsampling\n")
cat("CV AUC:", round(xgb_down_auc, 4), "\n")
cat("Sensitivity:", round(xgb_down$results$Sens, 4), "\n")
cat("Specificity:", round(xgb_down$results$Spec, 4), "\n")
```

Downsampling reduced the majority class to match the minority class size.

## 3.3 Upsampling

```{r model-upsample}
# Set up cross-validation with upsampling
train_control_up <- trainControl(
  method = "cv",
  number = cv_folds,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final",
  sampling = "up",
  verboseIter = FALSE
)

# Train XGBoost with upsampling (use complete data)
xgb_up <- train(
  TARGET ~ .,
  data = xgb_sample,
  method = "xgbTree",
  trControl = train_control_up,
  metric = "ROC",
  tuneGrid = xgb_best_grid,
  verbosity = 0
)

xgb_up_auc <- xgb_up$results$ROC

cat("XGBoost with Upsampling\n")
cat("CV AUC:", round(xgb_up_auc, 4), "\n")
cat("Sensitivity:", round(xgb_up$results$Sens, 4), "\n")
cat("Specificity:", round(xgb_up$results$Spec, 4), "\n")
```

Upsampling increased the minority class by duplicating observations.

## 3.4 SMOTE

```{r model-smote}
if (!requireNamespace("themis", quietly = TRUE)) {
  xgb_smote_auc <- NA_real_
  xgb_smote_sens <- NA_real_
  xgb_smote_spec <- NA_real_
  cat("Skipping SMOTE because package 'themis' is not installed.\n")
} else {
  # Use a smaller sample because synthetic oversampling is expensive.
  smote_sample <- safe_sample_n(model_data_complete, smote_sample_size)

  train_control_smote <- trainControl(
    method = "cv",
    number = cv_folds,
    classProbs = TRUE,
    summaryFunction = twoClassSummary,
    savePredictions = "final",
    sampling = "smote",
    verboseIter = FALSE
  )

  xgb_smote <- train(
    TARGET ~ .,
    data = smote_sample,
    method = "xgbTree",
    trControl = train_control_smote,
    metric = "ROC",
    tuneGrid = xgb_best_grid,
    verbosity = 0
  )

  xgb_smote_auc <- xgb_smote$results$ROC
  xgb_smote_sens <- xgb_smote$results$Sens
  xgb_smote_spec <- xgb_smote$results$Spec

  cat("XGBoost with SMOTE\n")
  cat("CV AUC:", round(xgb_smote_auc, 4), "\n")
  cat("Sensitivity:", round(xgb_smote_sens, 4), "\n")
  cat("Specificity:", round(xgb_smote_spec, 4), "\n")
}
```

SMOTE creates synthetic examples of the minority class rather than duplicating existing observations.

## 3.5 Class Imbalance Strategy Comparison

```{r imbalance-comparison}
# Create comparison table
imbalance_comparison <- tibble(
  Strategy = c("No Adjustment", "Downsampling", "Upsampling", "SMOTE"),
  CV_AUC = c(
    round(xgb_auc, 4),
    round(xgb_down_auc, 4),
    round(xgb_up_auc, 4),
    round(xgb_smote_auc, 4)
  ),
  Sensitivity = c(
    round(xgb_model$results$Sens, 4),
    round(xgb_down$results$Sens, 4),
    round(xgb_up$results$Sens, 4),
    round(xgb_smote_sens, 4)
  ),
  Specificity = c(
    round(xgb_model$results$Spec, 4),
    round(xgb_down$results$Spec, 4),
    round(xgb_up$results$Spec, 4),
    round(xgb_smote_spec, 4)
  )
) %>%
  arrange(desc(CV_AUC))

knitr::kable(imbalance_comparison,
             caption = "Class Imbalance Strategy Comparison (XGBoost)",
             col.names = c("Strategy", "CV AUC", "Sensitivity", "Specificity"))
```

The comparison shows trade-offs between detecting defaults (sensitivity) and correctly identifying non-defaults (specificity).


# 4. Supplementary Data

## 5.1 Load POS_Cash
```{r load-pos-cash}
# POS_CASH_balance contains monthly snapshots of point-of-sale and cash loans
# We aggregate to one row per SK_ID_CURR with summary statistics
pos_cash <- read_csv("Data/POS_CASH_balance.csv", show_col_types = FALSE)

pos_features <- pos_cash %>%
  group_by(SK_ID_CURR) %>%
  summarise(
    POS_MONTHS_BALANCE     = n(),                                        # total months of history
    POS_CNT_INSTALMENT_MEAN = mean(CNT_INSTALMENT, na.rm = TRUE),        # avg installments remaining
    POS_SK_DPD_MEAN        = mean(SK_DPD, na.rm = TRUE),                 # avg days past due
    POS_SK_DPD_MAX         = max(SK_DPD, na.rm = TRUE),                  # worst days past due
    POS_COMPLETED_RATIO    = mean(NAME_CONTRACT_STATUS == "Completed",   # share of completed contracts
                                  na.rm = TRUE),
    .groups = "drop"
  )

cat("POS_CASH features created:", ncol(pos_features) - 1, "features for", 
    nrow(pos_features), "customers\n")

```

## 5.2 Load Credit Card Balance
```{r load-cc-balance}
# credit_card_balance contains monthly credit card snapshots
# Key signals: utilization rate, payment behavior, drawing patterns
cc_balance <- read_csv("Data/credit_card_balance.csv", show_col_types = FALSE)

cc_features <- cc_balance %>%
  group_by(SK_ID_CURR) %>%
  summarise(
    CC_MONTHS_BALANCE      = n(),                                              # total months of history
    CC_AMT_BALANCE_MEAN    = mean(AMT_BALANCE, na.rm = TRUE),                  # avg balance
    CC_AMT_DRAWINGS_MEAN   = mean(AMT_DRAWINGS_CURRENT, na.rm = TRUE),         # avg monthly drawings
    CC_SK_DPD_MEAN         = mean(SK_DPD, na.rm = TRUE),                       # avg days past due
    CC_SK_DPD_MAX          = max(SK_DPD, na.rm = TRUE),                        # worst days past due
    CC_UTILIZATION_MEAN    = mean(AMT_BALANCE / (AMT_CREDIT_LIMIT_ACTUAL + 1), # avg utilization rate
                                  na.rm = TRUE),
    .groups = "drop"
  )

cat("Credit card features created:", ncol(cc_features) - 1, "features for",
    nrow(cc_features), "customers\n")

```

## 5.3 Join Data
```{r join-data}
# Since model_data_complete dropped SK_ID_CURR, we rebuild from model_data
# (which still has it) and re-apply na.omit after joining

model_data_with_id <- train_data %>%
  mutate(TARGET = factor(ifelse(TARGET == 1, "Default", "NoDefault"),
                         levels = c("NoDefault", "Default"))) %>%
  left_join(pos_features,  by = "SK_ID_CURR") %>%
  left_join(cc_features,   by = "SK_ID_CURR") %>%
  select(-SK_ID_CURR)

# Fill NAs from join (customers with no POS/CC history)
new_cols <- c(names(pos_features)[-1], names(cc_features)[-1])
model_data_with_id[new_cols] <- lapply(
  model_data_with_id[new_cols],
  function(x) replace(x, is.na(x), 0)
)

# Drop remaining NAs and single-value columns
model_data_enhanced <- model_data_with_id %>% na.omit()

model_data_enhanced_prepped <- prepare_model_frame(model_data_enhanced)
model_data_enhanced <- model_data_enhanced_prepped$data

cat("Enhanced dataset:", nrow(model_data_enhanced), "rows,", 
    ncol(model_data_enhanced) - 1, "features\n")
cat("New features added:", length(new_cols), "\n")

```

## 5.4 Train XGBoost with Enhanced Data
```{r model-xgb-enhanced}
## 5.4 Model Comparison: With vs Without POS/CC Features

# Train XGBoost on enhanced dataset using same fixed hyperparameters as Section 2
# for a fair apples-to-apples comparison
enhanced_sample <- safe_sample_n(model_data_enhanced, xgb_sample_size)

xgb_enhanced <- train(
  TARGET ~ .,
  data = enhanced_sample,
  method = "xgbTree",
  trControl = train_control,   # same 3-fold CV as before
  metric = "ROC",
  tuneGrid = xgb_best_grid,
  verbosity = 0
)

xgb_enhanced_auc <- xgb_enhanced$results$ROC

final_model <- xgb_enhanced

# Summary comparison table
supplementary_comparison <- tibble(
  Model = c(
    "XGBoost - App features only",
    "XGBoost - With bureau/prev/install",
    "XGBoost - With all supplementary (incl. POS/CC)"
  ),
  Features = c(
    "Application data only",
    "App + bureau + previous + installments",
    "App + bureau + previous + installments + POS + CC"
  ),
  CV_AUC = c(
    round(xgb_app_auc, 4),
    round(xgb_auc, 4),
    round(xgb_enhanced_auc, 4)
  )
)

knitr::kable(supplementary_comparison,
             caption = "Impact of Supplementary Data on XGBoost Performance",
             col.names = c("Model", "Features", "CV AUC"))

```



# 6. Kaggle Submission

## 6.1 Generate Submission
```{r generate-submission}
# Generate predicted probabilities for test set
test_pos_cash <- read_csv("Data/POS_CASH_balance.csv", show_col_types = FALSE) %>%
  group_by(SK_ID_CURR) %>%
  summarise(
    POS_MONTHS_BALANCE = n(),
    POS_CNT_INSTALMENT_MEAN = mean(CNT_INSTALMENT, na.rm = TRUE),
    POS_SK_DPD_MEAN = mean(SK_DPD, na.rm = TRUE),
    POS_SK_DPD_MAX = max(SK_DPD, na.rm = TRUE),
    POS_COMPLETED_RATIO = mean(NAME_CONTRACT_STATUS == "Completed", na.rm = TRUE),
    .groups = "drop"
  )

test_cc_balance <- read_csv("Data/credit_card_balance.csv", show_col_types = FALSE) %>%
  group_by(SK_ID_CURR) %>%
  summarise(
    CC_MONTHS_BALANCE = n(),
    CC_AMT_BALANCE_MEAN = mean(AMT_BALANCE, na.rm = TRUE),
    CC_AMT_DRAWINGS_MEAN = mean(AMT_DRAWINGS_CURRENT, na.rm = TRUE),
    CC_SK_DPD_MEAN = mean(SK_DPD, na.rm = TRUE),
    CC_SK_DPD_MAX = max(SK_DPD, na.rm = TRUE),
    CC_UTILIZATION_MEAN = mean(AMT_BALANCE / (AMT_CREDIT_LIMIT_ACTUAL + 1), na.rm = TRUE),
    .groups = "drop"
  )

test_model_enhanced <- test_data %>%
  left_join(test_pos_cash, by = "SK_ID_CURR") %>%
  left_join(test_cc_balance, by = "SK_ID_CURR")

test_model_enhanced[new_cols] <- lapply(
  test_model_enhanced[new_cols],
  function(x) replace(x, is.na(x), 0)
)

required_predictors <- setdiff(names(final_model$trainingData), ".outcome")
missing_predictors <- setdiff(required_predictors, names(test_model_enhanced))

if (length(missing_predictors) > 0) {
  stop("Test data is missing predictors required by final_model: ",
       paste(missing_predictors, collapse = ", "))
}

test_model_enhanced <- test_model_enhanced[, required_predictors, drop = FALSE]
test_model_enhanced <- test_model_enhanced %>%
  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))

if (length(final_model$xlevels) > 0) {
  for (col_name in names(final_model$xlevels)) {
    if (col_name %in% names(test_model_enhanced)) {
      valid_levels <- final_model$xlevels[[col_name]]
      coerced_values <- as.character(test_model_enhanced[[col_name]])
      coerced_values[is.na(coerced_values) | !(coerced_values %in% valid_levels)] <- valid_levels[1]
      test_model_enhanced[[col_name]] <- factor(coerced_values, levels = valid_levels)
    }
  }
}

test_preds <- predict(final_model, newdata = test_model_enhanced, type = "prob")

# Create submission file
submission <- tibble(
  SK_ID_CURR = test_data$SK_ID_CURR,
  TARGET = test_preds$Default
)

write_csv(submission, "submission.csv")
cat("Submission file created:", nrow(submission), "rows\n")
head(submission)
```

## 6.2 Kaggle Score
